---
title: "FIN 6155 · Digital Credit Build · Stage 1 starter — Plaid Sandbox pipeline"
author: "WageWave"
output: html_document
---


Plaid docs worth skimming: Sandbox <https://plaid.com/docs/sandbox/> · test
credentials & personas <https://plaid.com/docs/sandbox/test-credentials/> ·
`/transactions/sync` <https://plaid.com/docs/api/products/transactions/>.



```{r setup}


WORKING_DIR <- ".."
setwd(WORKING_DIR)                          # keys, raw/ downloads, outputs all live here
if (file.exists(".Renviron")) readRenviron(".Renviron")  # load PLAID_CLIENT_ID / PLAID_SECRET

library(httr2); library(jsonlite); library(tidyverse)

PLAID_HOST <- "https://sandbox.plaid.com"
CLIENT_ID  <- Sys.getenv("PLAID_CLIENT_ID")
SECRET     <- Sys.getenv("PLAID_SECRET")
stopifnot("Set PLAID_CLIENT_ID in your project .Renviron" = nzchar(CLIENT_ID),
          "Set PLAID_SECRET in your project .Renviron"    = nzchar(SECRET))

dir.create("raw", showWarnings = FALSE)   # all downloads land here

# Low-level helper: POST one Plaid endpoint, return the parsed JSON body.
# Read it top to bottom — each line is one step (keep the |> at line end):
plaid_post <- function(endpoint, body = list()) {
  request(paste0(PLAID_HOST, endpoint)) |>                               # 1. URL = host + endpoint
    req_body_json(c(list(client_id = CLIENT_ID, secret = SECRET), body)) |>  # 2. attach auth + this call's fields
    req_error(body = \(resp) resp_body_string(resp)) |>                  # 3. on HTTP error, surface Plaid's JSON message
    req_perform() |>                                                     # 4. send the request
    resp_body_json()                                                     # 5. parse JSON -> R list
}
```

## 1. Open one Item per persona (WORKED EXAMPLE)

In production a user logs into their bank through Plaid Link. In Sandbox you
skip Link: `/sandbox/public_token/create` mints an Item for a fake user at a
test institution, which you exchange for an `access_token`. We pin `ins_109508`
("First Platypus Bank") because the credit-profile personas only work at a
non-OAuth institution — it is required, not a stand-in for a real bank.



#Takes one persona and created an access_token to bank "ins_109508" to initialize transaction product and return access token for persona
```{r open-item}
#persona, bank
open_item <- function(persona, institution = "ins_109508") {
  tryCatch(
    {
      pt <- plaid_post("/sandbox/public_token/create", list(
        institution_id   = institution,
        initial_products = list("transactions"),
        options = list(override_username = persona)
      ))

      plaid_post(
        "/item/public_token/exchange",
        list(public_token = pt$public_token)
      )$access_token
    },

    error = function(e) {
      message("Failed to open item for ", persona)
      return(NULL)
    }
  )
}
```

## 2. Pull EVERYTHING for one customer, then keep what you need

Rather than scatter calls across your script, pull all products for one Item
into a single **bundle** — the same shape as the Stage 2 data release
(`{accounts, transactions, identity, liabilities}`). Your Stage 2 loader will
read bundles exactly like the one you persist here, so this is the same code.

```{r pull-customer}
sync_transactions <- function(access_token) {
  
  cursor <- ""
  added <- list()
  transactions_error <- NA_character_
  
  tryCatch(
    {
      repeat {
        resp <- plaid_post(
          "/transactions/sync",
          list(
            access_token = access_token,
            cursor = cursor,
            count = 500
          )
        )
        
        added <- c(added, resp$added)
        cursor <- resp$next_cursor
        
        if (!isTRUE(resp$has_more)) break
      }
      
    },
    error = function(e) {
      transactions_error <<- "CANNOT_SYNC_TRANSACTIONS"
      message("sync transactions error: ", conditionMessage(e))
    }
  )
  
  list(
    transactions = added,
    error = transactions_error
  )
}


pull_customer <- function(persona) {
  #access_token = tok
  
  tok <- open_item(persona)
  
  if (is.null(tok)) {
    return(list(
      bundle = NULL,
      item_error = "ITEM_CREATION_FAILED"
    ))
  }

  liab_error <- NA_character_
  accounts_error <- NA_character_
  identity_error <- NA_character_
  
  #.    *ADD similar edge cases for the other endpoints*
  
  #pull liabilities if there are any and add ti liab, otherwise save the error message to use in inventory later
  # Use the access token to request this customer's liability data
  liab <- tryCatch(plaid_post("/liabilities/get", list(access_token = tok))$liabilities,
                   error = function(e) {
                     #catches the error if there is one
                     liab_error <<- "NO_LIABILITY_ACCOUNTS"
                     message("liabilities error for ", persona, ": ", conditionMessage(e)
                             )
                     NULL
                   })
  accounts <- tryCatch(plaid_post("/accounts/get", list(access_token = tok))$accounts,
                   error = function(e) {
                     #catches the error if there is one
                     accounts_error <<- "NO_ACCOUNTS"
                     message("accounts error for ", persona, ": ", conditionMessage(e)
                             )
                     NULL
                   })
    identity <- tryCatch(plaid_post("/identity/get", list(access_token = tok))$accounts,
                   error = function(e) {
                     #catches the error if there is one
                     identity_error <<- "NO_IDENTITIES"
                     message("identities error for ", persona, ": ", conditionMessage(e)
                             )
                     NULL
                   })
    
  tx_result <- sync_transactions(tok)
    
  bundle <- list(
    customer_id  = persona,
    accounts     = accounts,
    identity     = identity,
    transactions = tx_result$transactions,
    liabilities  = liab
  )
  
  #creates raw directory
  dir.create("raw", showWarnings = FALSE, recursive = TRUE)
  
  write_json(bundle, sprintf("raw/%s.json", persona), auto_unbox = TRUE, null = "null")
  bundle
  
  list(
    bundle = bundle,
    liab_error = liab_error,
    accounts_error = accounts_error,
    identity_error = identity_error,
    transactions_error = tx_result$error
)
}

#b <- pull_customer("user_credit_profile_good") #This takes one persona "user_credit_profile_good" and creates the transactions JSON
#str(b, max.level = 1)


```

## 3. Inspect it as a table / CSV (optional — for your eyes)

```{r flatten}
# Flatten one bundle's transactions to a tidy tibble (and CSV) so you can open
# them in a spreadsheet. This is also where your Stage 2 feature work begins.
flatten_tx <- function(b) {
  map_dfr(b$transactions, \(t) tibble(
    date          = t$date,
    amount        = t$amount,                                   # mind the Plaid sign convention!
    name          = t$name %||% NA_character_,
    merchant_name = t$merchant_name %||% NA_character_,
    pfc_primary   = t$personal_finance_category$primary  %||% NA_character_,
    pfc_detailed  = t$personal_finance_category$detailed %||% NA_character_
  ))
}
#tx <- flatten_tx(b$bundle)
#head(tx)

#In Plaid's transaction convention, negative amounts represent inflows (money entering the account), while positive amounts represent outflows (money leaving the account).

```

## 4. Scale out — YOUR TURN


```{r extend}
#the 4 personas
user_personas <- c("user_credit_profile_poor", "user_credit_profile_good", "user_credit_profile_excellent", "user_credit_bonus")

pull_multiple_customers <- function(personas) {
  results <- list()
  inventory <- tibble()
  for (persona in personas) {
    results[[persona]] <- pull_customer(persona)
    
    if (is.null(results[[persona]]$bundle)) {
      inventory <- bind_rows(
        inventory,
        tibble(
          persona = persona,
          endpoint = "item_creation",
          n_records = 0,
          date_min = NA_character_,
          date_max = NA_character_,
          error = results[[persona]]$item_error
        )
      )
      
      next
    }
    
    customer <-results[[persona]]$bundle
    liab_error <- results[[persona]]$liab_error
    accounts_error <- results[[persona]]$accounts_error 
    identity_error <- results[[persona]]$identity_error
    transactions_error <- results[[persona]]$transactions_error
    
    
    # ACCOUNTS
    account_row <- tibble(
      persona = persona,
      endpoint = "/accounts/get",
      n_records = length(customer$accounts),
      date_min = NA_character_,
      date_max = NA_character_,
      error = accounts_error
    )
    
    
    # IDENTITY
    identity_row <- tibble(
      persona = persona,
      endpoint = "/identity/get",
      n_records = length(customer$identity),
      date_min = NA_character_,
      date_max = NA_character_,
      error = identity_error
    )
    
    
    # TRANSACTIONS
    tx <- flatten_tx(customer)
    
    transaction_row <- tibble(
      persona = persona,
      endpoint = "/transactions/sync",
      n_records = nrow(tx),
      date_min = if (nrow(tx) > 0) min(tx$date, na.rm = TRUE) else NA_character_,
      date_max = if (nrow(tx) > 0) max(tx$date, na.rm = TRUE) else NA_character_,
      error = transactions_error
    )
    
    
    # LIABILITIES
    liability_row <- tibble(
      persona = persona,
      endpoint = "/liabilities/get",
      n_records = length(customer$liabilities),
      date_min = NA_character_,
      date_max = NA_character_,
      error = liab_error
    )
    
    inventory <- bind_rows(
      inventory,
      account_row,
      identity_row,
      transaction_row,
      liability_row
    )
    
  }
  write_csv(inventory, "raw/raw_data_inventory.csv")
  list(
    results = results,
    inventory = inventory
  )
  
}

all_customers <- pull_multiple_customers(user_personas)



```

## 5. Reflection (graded — see Handout, Stage 1)

In <= 1 page, citing your inventory: (i) which endpoints your product consumes
and why; (ii) what Section 1033 does to your cost of access to this data;
(iii) why a handful of personas cannot train a credit model — what exactly is
missing statistically?
