---
title: "FIN 6155 · Digital Credit Build · Stage 2 starter — synthetic data loader + baseline model"
author: "(your team)"
output: html_document
---

You receive a historical loan book from "First Ute Bank" — 25,000 applicants
in your team's release (20,000 historical + 5,000 pending) — as
Plaid-shaped JSON (a folder of 50 shards, 500 customers each) plus an applications file. It is EQUILIBRIUM data —
an incumbent lender APPROVED some applicants (`applications.csv$approved`), PRICED
them (`offered_apr`), and only some ACCEPTED (`took_up`). Performance
(`default_12m` in `outcomes_train.csv`) is observed ONLY for taken-up loans —
roughly 7,200–9,300 rows, the exact count depending on your cohort. The
5,000 ids in `holdout_ids.csv` are pending applicants (no funnel, no outcome) that you
underwrite and price; your predictions are scored against the instructor answer key.

Because labels exist only for taken-up loans, your PD model is trained on a
doubly-selected sample and will be MISCALIBRATED on the full applicant population.
And `offered_apr` was never randomized, so the price–default correlation here is
selection, NOT the causal effect of price — you cannot recover how behaviour
changes when prices change. Reason about both in your write-up.

The transaction records have EXACTLY the schema your Stage 1 pipeline produced.
Your Stage 1 parsing knowledge transfers one-for-one.

```{r setup}
# ===== SET YOUR WORKING DIRECTORY FIRST (everything saves/loads here) =====
# Leave "." if you opened digital-credit-build.Rproj; else the folder's full path.
WORKING_DIR <- "."
setwd(WORKING_DIR)                          # keys, raw/ downloads, outputs all live here
if (file.exists(".Renviron")) readRenviron(".Renviron")  # load PLAID_CLIENT_ID / PLAID_SECRET

library(tidyverse)
library(jsonlite)
# Modeling stack (Stage 2 baseline uses plain glm; tidymodels allowed):
# install.packages(c("tidymodels", "glmnet", "ranger", "yardstick"))

DATA_DIR <- "data_release"   # adjust to where you unpacked the release
```

## 1. Load one shard (WORKED EXAMPLE — then scale up)

Each shard is gzipped NDJSON: one customer bundle per line.

```{r load-one}
read_shard <- function(path) {
  con <- gzfile(path, "rt")
  on.exit(close(con))
  lines <- readLines(con, warn = FALSE)
  lapply(lines, fromJSON, simplifyVector = TRUE)
}

shard0 <- read_shard(file.path(DATA_DIR, "plaid_synth_shard_00.ndjson.gz"))
length(shard0)              # 500 customers
b <- shard0[[1]]            # one customer bundle
str(b, max.level = 1)       # customer_id, item, identity, accounts, transactions, liabilities
head(as_tibble(b$transactions))
```

## 2. Feature engineering (WORKED EXAMPLE: 2 features; you need many more)

Remember the Plaid sign convention you documented in Stage 1.

```{r features-example}
customer_features <- function(b) {
  tx <- as_tibble(b$transactions) |>
    mutate(date = as.Date(date),
           pfc_primary  = personal_finance_category$primary,
           pfc_detailed = personal_finance_category$detailed)

  # Example feature 1: mean monthly income, identified BY CATEGORY
  inc_m <- tx |>
    filter(pfc_primary == "INCOME", amount < 0) |>
    mutate(month = format(date, "%Y-%m")) |>
    group_by(month) |> summarise(inc = -sum(amount), .groups = "drop")
  inc_mean <- mean(inc_m$inc)

  # Example feature 2: overdraft (NSF) events
  nsf <- sum(tx$pfc_detailed == "BANK_FEES_OVERDRAFT_FEES")

  tibble(customer_id = b$customer_id, inc_mean = inc_mean, nsf = nsf)

  # TODO — ideas you should pursue (non-exhaustive; discretion is graded):
  #  * income VOLATILITY (cv of monthly income) and payroll regularity
  #  * balance path: you observe the END balance in b$accounts and every dated
  #    flow — reconstruct the daily balance BACKWARD from the end balance.
  #    Median balance / income ("buffer") is a strong cash-flow feature.
  #  * spending composition; debt payments (LOAN_PAYMENTS) / income
  #  * credit card utilization from the credit-type account (balance / limit);
  #    liabilities: minimum_payment_amount, is_overdue (watch missing cases!)
  #  * data quality: are any transactions duplicated? do all "income-looking"
  #    inflows belong in income? (TRANSFER_IN is your own money moving.)
  #  * application file: requested_amount vs income; STATED income vs what the
  #    bank data shows (W2: "Observing Unobservables"...)
}

customer_features(b)
```

## 3. Scale to all shards + join labels

```{r build-table}
# TODO: loop all shards 00..49 (this takes a few minutes; cache the result!)
# files <- list.files(DATA_DIR, pattern = "^plaid_synth_shard_.*\\.ndjson\\.gz$",
#                     full.names = TRUE)
# X <- map_dfr(files, \(f) map_dfr(read_shard(f), customer_features))
# write_rds(X, "features_cache.rds")

apps   <- read_csv(file.path(DATA_DIR, "applications.csv"))
ytrain <- read_csv(file.path(DATA_DIR, "outcomes_train.csv"))
# df <- X |> left_join(apps, by = "customer_id") |>
#            left_join(ytrain, by = "customer_id")
```

IMPORTANT — labels (`outcomes_train.csv`) exist ONLY for taken-up loans (approved
AND accepted); a PD model fit on them is miscalibrated on all applicants. The
project asks for three models: an underwriting model (approve/deny, on all
applicants), a pricing model (APR for those you approve), and a PD model (on
taken-up loans). Apply them to the new-consumer holdout.

IMPORTANT — `group` in `applications.csv` is fair-lending MONITORING data
(think HMDA). It must NOT be used as a model input at any stage. You will use
it in Stage 3 to audit your model. Document this exclusion in your write-up.

## 4. Baseline logistic regression (Stage 2 deliverable)

```{r baseline}
# Split the LABELED taken-up customers (the rows of YOUR outcomes_train.csv;
# roughly 7,200-9,300, cohort-specific) into your own train/validation sets.
# (The pending ids in holdout_ids.csv are scored by the instructor.)
# set.seed(6155)
# TODO: glm(default_12m ~ ..., family = binomial(), data = train)
# TODO: validation AUC + KS (yardstick::roc_auc; KS from cumulative separation)
# TODO: calibration table: 10 deciles of predicted PD vs observed default rate
# TODO: coefficient table with one-sentence ECONOMIC interpretation per feature
#       (why SHOULD this predict default? "it helps AUC" is not an explanation)
```

## 5. Holdout submission (required with EVERY model stage)

```{r submission}
# TODO: predict PD for every id in holdout_ids.csv and write:
#   write_csv(tibble(customer_id = ..., pd = ...),
#             "TEAMNAME_group_NN_stage2.csv")   # NN = your data-group number
# Constraints: every pending id exactly once; pd strictly in (0, 1).
# Lowest-PD 70% of applicants would be approved under the default policy —
# but cutoff/pricing analysis is yours to design in Stages 3 and 5.
```

<!-- pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad -->
<!-- pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad pad -->
